Based on polling as of 27 October 2024, Harris expects to win on xxx%, leads Trump by xx% in 2024 US election

### 2024-10-30 paper/paper.qmd
merge conflict

```{r}
# Set seed for reproducibility
set.seed(123)

# Sample indices for training and testing sets
train_index <- sample(seq_len(nrow(data)), size = 0.8 * nrow(data))

# Split the data into training and testing sets
train_data <- data[train_index, ]
test_data <- data[-train_index, ]

# Fit the weighted linear regression model
first_model_test <- lm(pct ~ national_poll + pollster + population, 
                     data = data, 
                  data = train_data)

# Predict on test data
predictions <- predict(first_model_test, newdata = test_data)

# Calculate RMSE using the correct outcome variable
# Assuming 'pct' is the actual observed outcome variable in test_data
rmse <- sqrt(mean((predictions - test_data$pct)^2, na.rm = TRUE))

# Print RMSE
print(paste("RMSE:", rmse))

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2
#residual plot
plot(first_model, which = 1,main ="Residual Plot of unweighted linear regression model residualsmodel" )
# Q-Q plot for the residuals of the model
qqnorm(residuals(first_model),main= "Q-Q Plot of unweighted linear regression model residuals
")
qqline(residuals(first_model), col = "red")

>>>>

```{r}
#| include: false
#### Calculate Weights ####
#harris_data <-read.csv("~/2024 US Election Forecast/data/02-analysis_data/analysis_data_Harris.csv")
#trump_data <- read.csv("~/2024 US Election Forecast/data/02-analysis_data/analysis_data_Trump.csv")
## if you haven't run the model script, use this code
# Calculate weights for Harris data
# harris_data <- harris_data %>%
#   mutate(
#     sample_size_weight = pmin(sample_size / 2300, 1),  # Cap weight at 1
#     pollster_quality_weight = numeric_grade / 4  # Scale numeric grade to [0, 1]
#   ) %>%
#   group_by(pollster) %>%
#   mutate(
#     recent_poll_count = sum(recency < 30),  # Count recent polls within the last 30 days
#     poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)  # Adjust frequency weight
#   ) %>%
#   ungroup() %>%
#   mutate(combined_weight = recency * sample_size_weight * poll_frequency_weight * pollster_quality_weight)  # Combine weights

# Calculate weights for Trump data (similar to Harris)
# trump_data <- trump_data %>%
#   mutate(
#     sample_size_weight = pmin(sample_size / 2300, 1),  # Cap weight at 1
#     pollster_quality_weight = numeric_grade / 4  # Scale numeric grade to [0, 1]
#   ) %>%
#   group_by(pollster) %>%
#   mutate(
#     recent_poll_count = sum(recency < 30),  # Count recent polls within the last 30 days
#     poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)  # Adjust frequency weight
#   ) %>%
#   ungroup() %>%
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#### Set Seed for Reproducibility ####
set.seed(567)

# Sample indices for training and testing sets for Harris data
train_index_harris <- sample(seq_len(nrow(harris_data)), size = 0.8 * nrow(harris_data))

# Split the data into training and testing sets
train_data_harris <- harris_data[train_index_harris, ]
test_data_harris <- harris_data[-train_index_harris, ]

# Convert pollster to factor in the training dataset
train_data_harris$pollster <- factor(train_data_harris$pollster)

# Get levels from the training dataset
pollster_levels <- levels(train_data_harris$pollster)

# Set the same levels for pollster in testing data
test_data_harris$pollster <- factor(test_data_harris$pollster, levels = pollster_levels)

#### Fit the Unweighted Linear Regression Model for Harris ####
model_harris_unweighted <- lm(pct ~ national_poll + pollster + population, 
                             data = train_data_harris)

# Predict on test data for unweighted model
predictions_unweighted <- predict(model_harris_unweighted, newdata = test_data_harris)

# Calculate RMSE for unweighted model
rmse_unweighted <- sqrt(mean((predictions_unweighted - test_data_harris$pct)^2, na.rm = TRUE))
cat("Harris Unweighted Model RMSE:", rmse_unweighted, "\n")

#### Fit the Weighted Linear Regression Model for Harris ####
model_harris_weighted <- lm(pct ~ national_poll + pollster + population, 
                           data = train_data_harris, weights = combined_weight)

# Predict on test data for weighted model
predictions_weighted <- predict(model_harris_weighted, newdata = test_data_harris)

# Calculate RMSE for weighted model
rmse_weighted <- sqrt(mean((predictions_weighted - test_data_harris$pct)^2, na.rm = TRUE))
cat("Harris Weighted Model RMSE:", rmse_weighted, "\n")
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#### Set Seed for Reproducibility ####
set.seed(123)

# Sample indices for training and testing sets for Trump data
train_index_trump <- sample(seq_len(nrow(trump_data)), size = 0.8 * nrow(trump_data))

# Split the data into training and testing sets
train_data_trump <- trump_data[train_index_trump, ]
test_data_trump <- trump_data[-train_index_trump, ]

# Convert pollster to factor in the training dataset
train_data_trump$pollster <- factor(train_data_trump$pollster)

# Get levels from the training dataset
pollster_levels <- levels(train_data_trump$pollster)

# Set the same levels for pollster in testing data
test_data_trump$pollster <- factor(test_data_trump$pollster, levels = pollster_levels)

#### Fit the Unweighted Linear Regression Model for Trump ####
model_trump_unweighted <- lm(pct ~ national_poll + pollster + population, 
                             data = train_data_trump)

# Predict on test data for unweighted model
predictions_unweighted <- predict(model_trump_unweighted, newdata = test_data_trump)

# Calculate RMSE for unweighted model
rmse_unweighted <- sqrt(mean((predictions_unweighted - test_data_trump$pct)^2, na.rm = TRUE))
cat("Trump Unweighted Model RMSE:", rmse_unweighted, "\n")

#### Fit the Weighted Linear Regression Model for Trump ####
model_trump_weighted <- lm(pct ~ national_poll + pollster + population, 
                           data = train_data_trump, weights = combined_weight)

# Predict on test data for weighted model
predictions_weighted <- predict(model_trump_weighted, newdata = test_data_trump)

# Calculate RMSE for weighted model
rmse_weighted <- sqrt(mean((predictions_weighted - test_data_trump$pct)^2, na.rm = TRUE))
cat("Trump Weighted Model RMSE:", rmse_weighted, "\n")
```

The RMSE (Root Mean Square Error) values for the models indicate how
well the models predict the percentage of support for Kamala Harris and
Donald Trump in the 2024 U.S. presidential election. For the Harris
models, the unweighted RMSE is approximately 2.24, while the weighted
model achieves a slightly better performance with an RMSE of about 2.19.
This improvement in the weighted model suggests that incorporating
weights—likely reflecting the reliability or relevance of the
polls—provides a more accurate prediction of Harris's support levels
compared to the unweighted approach. The smaller RMSE value signifies
that the predictions made by the weighted model are closer to the actual
observed percentages, indicating better model performance.

In contrast, the RMSE values for Trump indicate a different scenario.
The unweighted model for Trump has an RMSE of approximately 2.97, while
the weighted model has a higher RMSE of about 3.07. This unexpected
outcome suggests that adding weights to the Trump model may have
introduced more variability or noise in the predictions rather than
improving accuracy. The higher RMSE in the weighted model compared to
the unweighted model indicates that the weighting scheme used might not
have been effective in capturing the true relationship between the
predictors and Trump’s support levels, potentially due to factors such
as multicollinearity or insufficient variation in the data. Overall,
while the Harris models show consistent improvement with weighting, the
Trump models demonstrate the need for further investigation into the
model specifications and the effectiveness of the weighting applied.

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Diagnostics of model using residual vs fitted plot and norm Q-Q plot\n-Support for Harris"
#| fig-subcap: ["Residual Plot for Unweighted Model", "Q-Q Plot for Unweighted Model", "Residual Plot for Weighted Model", "Q-Q Plot for Weighted Model"]
#| layout-ncol: 2

# Residual plot for unweighted model
plot(harris_unweighted_model, which = 1, main = "Residual Plot of Unweighted Linear Regression Model")

# Q-Q plot for the residuals of the unweighted model
qqnorm(residuals(harris_unweighted_model), main = "Q-Q Plot of Unweighted Linear Regression Model Residuals")
qqline(residuals(harris_unweighted_model), col = "red")

# Residual plot for weighted model
plot(harris_weighted_model, which = 1, main = "Residual Plot of Weighted Linear Regression Model")

# Q-Q plot for the residuals of the weighted model
qqnorm(residuals(harris_weighted_model), main = "Q-Q Plot of Weighted Linear Regression Model Residuals")
qqline(residuals(harris_weighted_model), col = "red")

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Diagnostics of model using residual vs fitted plot and norm Q-Q plot\n-Support for Trump"
#| fig-subcap: ["Residual Plot for Unweighted Model", "Q-Q Plot for Unweighted Model", "Residual Plot for Weighted Model", "Q-Q Plot for Weighted Model"]
#| layout-ncol: 2

# Residual plot for unweighted model
plot(trump_unweighted_model, which = 1, main = "Residual Plot of Unweighted Linear Regression Model")

# Q-Q plot for the residuals of the unweighted model
qqnorm(residuals(trump_unweighted_model), main = "Q-Q Plot of Unweighted Linear Regression Model Residuals")
qqline(residuals(trump_unweighted_model), col = "red")

# Residual plot for weighted model
plot(trump_weighted_model, which = 1, main = "Residual Plot of Weighted Linear Regression Model")

# Q-Q plot for the residuals of the weighted model
qqnorm(residuals(trump_weighted_model), main = "Q-Q Plot of Weighted Linear Regression Model Residuals")
qqline(residuals(trump_weighted_model), col = "red")
